---
title: "Penguin flipper length and body mass"
author: "Cesar Figueroa"
date: 2026-09-24
engine: knitr
---

## Question

Do penguins with longer flippers tend to have greater body mass? You wouldn't be surprise once you know the answer. 

## Data source

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/),
collected by Dr. Kristen Gorman and Palmer Station Antarctica LTER.
The data are available under a CC0 license.

## Load the data

```{r}
#| label: load-penguins
#| message: false

library(palmerpenguins)
head(penguins)
```

## Check data quality

I check for missing values in flipper length and body mass before
examining their relationship.

```{r}
#| label: check-missing-values

measurements <- penguins[c("flipper_length_mm", "body_mass_g")]

nrow(measurements)
colSums(is.na(measurements))
```

## Prepare the data

I retain observations with both flipper length and body mass available.
Missing values in other variables do not affect this selection.

```{r}
#| label: remove-missing-measurements

measurements_clean <- measurements[complete.cases(measurements), ]

nrow(measurements_clean)
nrow(measurements) - nrow(measurements_clean)
```

## Visualize the relationship

Each point represents one penguin with both measurements available.

```{r}
#| label: fig-flipper-mass
#| fig-cap: "Flipper length and body mass for 342 penguins with complete measurements."

plot(
  measurements_clean$flipper_length_mm,
  measurements_clean$body_mass_g,
  xlab = "Flipper length (mm)",
  ylab = "Body mass (g)",
  pch = 19,
  col = "steelblue"
)
```

## Measure the association

I calculate Pearson's correlation to summarize the linear association
between flipper length and body mass.

```{r}
#| label: calculate-correlation

correlation <- cor(
  measurements_clean$flipper_length_mm,
  measurements_clean$body_mass_g
)

round(correlation, 3)
```

## Findings :O

Among the 342 penguins with both measurements available, flipper
length and body mass have a Pearson correlation of 0.871. This
indicates a strong positive linear association, consistent with
the upward pattern in @fig-flipper-mass.

This analysis combines three species. Differences between species
may contribute to the overall association, so the result does not
necessarily describe the relationship within each species.
Correlation does not establish that longer flippers cause greater
body mass. Two observations with missing measurements were excluded.